### 2021-04-09 end_appendix/appendix.tex
Spell- and grammer-checked remaining chapters

The basic elements of the \ac{GBRT} algorithm are decision trees. These models
create decision rules based on binary splits to predict a target variable $y$
from a set of predictors $\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)}
\right)$ ($K$ is the number of predictors). These predictors do not need to be
of the same type: \ac{GBRT} allows the simultaneous input of numerical and
categorical variables, which is a great advantage for our use case. There is no
need to encode the categorical predictors in any way. Due to their simple
nature, \acf{ML} models based on decision trees are easy to interpret and
explain but cannot be used to create satisfying predictions for complex
datasets. This issue can be overcome by a technique called \emph{boosting}.
Boosting improves the performance of \emph{weak learners} (in our case decision
trees) by combining a large number of them  \autocite{Freund1996}. The
regression function used to predict $\hat{y} = F(\bm{x})$ can be written as a
linear combination of simple decision trees $h(\bm{x})$
\begin{equation}
  F(\bm{x}) = \sum_{j=0}^{H} \beta_j h(\bm{x}, \bm{\alpha}_j),
  \label{eq:app:b:f_x_sum}
\end{equation}
where $H$ is the total number of decision trees, $\beta_j$ the expansion
coefficient for tree $j$, and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid i
\in I_N \right\}$ with index set $I_N = \left\{ 1, 2, \ldots, N \right\}$
(\emph{classic gradient boosting}), the expansion coefficients and tree
parameters are jointly fitted by minimizing a loss function $L \left( y,
F(\bm{x}) \right)$ in a forward iteration:
\begin{equation}
  \left( \beta_j, \bm{\alpha}_j \right) = \underset{\beta,
    \bm{\alpha}}{\arg\min} \sum_{i=1}^{N} L \left( y_i, F_{j-1}(\bm{x}_i) +
  \beta h(\bm{x}_i, \bm{\alpha}) \right).
  \label{eq:app:b:beta_alpha}
\end{equation}
In practice, this iteration step only uses a randomly selected subsample of the
training data (drawn without replacement), \ie{} the sum does not cover all $N$
training points. Starting with an initial guess $F_0(\bm{x})$, the model is
recursively built by
\begin{equation}
  F_j(\bm{x}) = F_{j-1}(\bm{x}) + \beta_j h(\bm{x}, \bm{\alpha}_j).
  \label{eq:app:b:f_x_recursion}
\end{equation}
The minimization procedure of the loss function (mean square error with
additional sample weights determined by the grid cell areas) is called
\emph{stochastic gradient boosting} and is explained in detail by
\textcite{Friedman2001, Friedman2002}. Fitting the \ac{GBRT} model involves
building the decision trees by splitting the data at points with maximum
information gain. Boosting those simple trees greatly improves the overall
predictive power of the \ac{ML} algorithm: Poorly modeled training points in
the early stages of the algorithm will gradually improve throughout the
training process.

A crucial criterion for the successful application of any \ac{GBRT} algorithm
is the choice of several hyperparameters. The three main control parameters of
the learning procedure are the total number of decision trees $H$, the
complexity of the individual trees (for example measured by the maximum tree
depth), and the learning rate $\nu \ll 1$. The latter parameter is used for
regularization and dramatically reduces the risk of overfitting by scaling down
the contribution of each added weak learner \autocite{Death2007, Elith2008,
  Friedman2001}. A common way to optimize the algorithm is $n$-fold \acf{CV}
\autocite{Bishop2006}: The data is randomly divided into a training and a
validation dataset and the \acs{GBRT} model is fitted on the training data
only. After that, the performance of this model can be evaluated on the
validation dataset by a suitable metric (\eg{} the mean square error). This
process is repeated $n$ times so that every input point is part of the
validation set at least once. The optimal hyperparameters are the
hyperparameters that show optimal performance on the validation datasets
\autocite{Elith2008}.

@@

The final source of uncertainty is the error of the prediction input data
$\sigma_{ik}$ ($i$ corresponds to the prediction input point and $k$ to the
predictor). These are only available for the FLUXNET-MTE product
\autocite{Jung2011}. To account for this, we use the \acf{LIME} technique
\commentcite{Ribeiro2016}{see \cref{subsec:06:step_2}} to build a local linear
model for every sample point, which yields the linear coefficients $b_{ik}$.
Using error propagation for all predictors and assuming independence of all
individual errors, the \ac{SPE} due to observational uncertainty $\sigma_{i,
  \mathtxt{OBS}}$ then be calculated as
\begin{equation}
  \sigma_{i, \mathtxt{OBS}} = \sum_{k=1}^{K} b_{ik}^2 \sigma_{ik}^2.
  \label{eq:app:b:sigma_obs}
\end{equation}
The total \ac{SPE} at a prediction input point $i$ is the sum of the squared
errors presented above (assuming all of them are independent):
\begin{equation}
  \sigma_i^2 = \sigma_\mathtxt{GBRT}^2 + \sigma_{i, \mathtxt{RESC}}^2 +
  \sigma_{i, \mathtxt{OBS}}^2.
  \label{eq:app:b:sigma_sum}
\end{equation}
The specified error ranges for the \ac{MMM} approaches are calculated in a
similar way: The constant \ac{SPE} per grid cell is estimated by the mean
\ac{RMSEP} given in the leave-one-model-out \ac{CV} experiment (see
\cref{subsec:02:model_weighting}). For the plain \ac{MMM}, this is the only
source of uncertainty. For the rescaled \ac{MMM}, the total error can be
calculated similarly to \cref{eq:app:b:sigma_sum} without the last term
(observational uncertainty).

@@

A convenient way to gain information about statistical models is to analyze the
residuals $\varepsilon_i$, which are defined as the difference between the true
value of the target variable $y_i$ and the predicted value $\hat{y}_i$ at a
sample point $i$ with known ground truth:
\begin{equation}
  \varepsilon_i = y_i - \hat{y}_i.
  \label{eq:app:b:residuals}
\end{equation}
A common way to visualize the residuals is to plot their \acfp{PDF} (see
\cref{fig:app:b:residuals}). The two panels (for steps 2a and 2b) show that the
\ac{ML} model is not overfitting in both cases: The distributions of the
training data and the independent hold-out data are very similar. Moreover, the
\acp{PDF} do not show significant biases, as the residuals are approximately
unbiased (zero mean) for the training and the test dataset. This justifies the
use of the \ac{RMSEP} to estimate the \ac{SPE}, since for unbiased residuals
the \ac{RMSEP} is equal to the standard deviation of the residuals.

@@

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4a.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4b.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:b}
  \end{subfigure}
  \\
    \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4c.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4d.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:d}
  \end{subfigure}
  \\
    \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4e.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4f.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:f}
  \end{subfigure}
  \caption[
    Geographical distributions of the absolute \acf{GPP} at the end of the
    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
    statistical models.
  ]{
    Geographical distributions of the absolute \acf{GPP} at the end of the
    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
    statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
    \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO} model using only the
    historical \acs{GPP} as single predictor. (d) \Acf{GBRT} model using only
    the historical \acs{GPP} as single predictor. (e) \acs{LASSO} model using
    all predictors. (f) \acs{GBRT} model using all predictors.
    \AdaptedFrom{Schlund2020}.
  }
  \label{fig:app:b:step2a_results}
\end{figure}

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5a.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5b.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5c.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5d.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:d}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5e.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5f.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:f}
  \end{subfigure}
  \caption[
    Geographical distributions of the \aclp{SPE} of the absolute \acf{GPP} at
    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
    different statistical models.
  ]{
    Geographical distributions of the \acfp{SPE} of the absolute \acf{GPP} at
    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
    different statistical models. Details on the calculation of the \acs{SPE}
    are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5 \acf{MMM}.
    (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO}
    model using only the historical \acs{GPP} as single predictor. (d)
    \Acf{GBRT} model using only the historical \acs{GPP} as single predictor.
    (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT} model using all
    predictors. The \acs{SPE} is minimal for the \acs{GBRT} model using all
    predictors. \AdaptedFrom{Schlund2020}.
  }
  \label{fig:app:b:step2a_results_errors}
\end{figure}

>>>>

The basic elements of the \ac{GBRT} algorithm are decision trees. These models
create decision rules based on binary splits to predict a target variable $y$
from a set of predictors $\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)}
\right)$ ($K$ is the number of predictors). These predictors do not need to be
of the same type: \ac{GBRT} allows the simultaneous input of numerical and
categorical variables, which is a great advantage for our use case. There is no
need to encode the categorical predictors in any way. Due to their simple
nature, \acf{ML} models based on decision trees are easy to interpret and
explain but cannot be used to create satisfying predictions for complex
datasets. This issue can be overcome by a technique called \emph{boosting}.
Boosting improves the performance of \emph{weak learners} (in our case decision
trees) by combining a large number of them  \autocite{Freund1996}. The
regression function used to predict $\hat{y} = F(\bm{x})$ can be written as a
linear combination of simple decision trees $h(\bm{x})$
\begin{equation}
  F(\bm{x}) = \sum_{j=0}^{H} \beta_j h(\bm{x}, \bm{\alpha}_j),
  \label{eq:app:b:f_x_sum}
\end{equation}
where $H$ is the total number of decision trees, $\beta_j$ the expansion
coefficient for tree $j$, and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid i
\in I_N \right\}$ with index set $I_N = \left\{ 1, 2, \ldots, N \right\}$
(\emph{classic gradient boosting}), the expansion coefficients and tree
parameters are jointly fitted by minimizing a loss function $L \left( y,
F(\bm{x}) \right)$ in a forward iteration:
\begin{equation}
  \left( \beta_j, \bm{\alpha}_j \right) = \underset{\beta,
    \bm{\alpha}}{\arg\min} \sum_{i=1}^{N} L \left( y_i, F_{j-1}(\bm{x}_i) +
  \beta h(\bm{x}_i, \bm{\alpha}) \right).
  \label{eq:app:b:beta_alpha}
\end{equation}
In practice, this iteration step only uses a randomly selected subsample of the
training data (drawn without replacement), \ie{} the sum does not cover all $N$
training points. Starting with an initial guess $F_0(\bm{x})$, the model is
recursively built by
\begin{equation}
  F_j(\bm{x}) = F_{j-1}(\bm{x}) + \beta_j h(\bm{x}, \bm{\alpha}_j).
  \label{eq:app:b:f_x_recursion}
\end{equation}
The minimization procedure of the loss function (mean square error with
additional sample weights determined by the grid-cell areas) is called
\emph{stochastic gradient boosting} and is explained in detail by
\textcite{Friedman2001, Friedman2002}. Fitting the \ac{GBRT} model involves
building the decision trees by splitting the data at points with maximum
information gain. Boosting those simple trees greatly improves the overall
predictive power of the \ac{ML} algorithm: Poorly modeled training points in
the early stages of the algorithm will gradually improve throughout the
training process.

A crucial criterion for the successful application of any \ac{GBRT} algorithm
is the choice of several hyperparameters. The three main control parameters of
the learning procedure are the total number of decision trees $H$, the
complexity of the individual trees (for example, measured by the maximum tree
depth), and the learning rate $\nu \ll 1$. The latter parameter is used for
regularization and dramatically reduces the risk of overfitting by scaling down
the contribution of each added weak learner \autocite{Death2007, Elith2008,
  Friedman2001}. A common way to optimize the algorithm is $n$-fold \acf{CV}
\autocite{Bishop2006}: The data is randomly divided into a training and a
validation dataset and the \acs{GBRT} model is fitted on the training data
only. After that, the performance of this model can be evaluated on the
validation dataset by a suitable metric (\eg{} the mean square error). This
process is repeated $n$ times so that every input point is part of the
validation set at least once. The optimal hyperparameters are the
hyperparameters that show optimal performance on the validation datasets
\autocite{Elith2008}.

@@

The final source of uncertainty is the error of the prediction input data
$\sigma_{ik}$ ($i$ corresponds to the prediction input point and $k$ to the
predictor). These are only available for the FLUXNET-MTE product
\autocite{Jung2011}. To account for this, we use the \acf{LIME} technique
\commentcite{Ribeiro2016}{see \cref{subsec:06:step_2}} to build a local linear
model for every sample point, which yields the linear coefficients $b_{ik}$.
Using error propagation for all predictors and assuming independence of all
individual errors, the \ac{SPE} due to observational uncertainty $\sigma_{i,
  \mathtxt{OBS}}$ then be calculated as
\begin{equation}
  \sigma_{i, \mathtxt{OBS}} = \sum_{k=1}^{K} b_{ik}^2 \sigma_{ik}^2.
  \label{eq:app:b:sigma_obs}
\end{equation}
The total \ac{SPE} at a prediction input point $i$ is the sum of the squared
errors presented above (assuming all of them are independent):
\begin{equation}
  \sigma_i^2 = \sigma_\mathtxt{GBRT}^2 + \sigma_{i, \mathtxt{RESC}}^2 +
  \sigma_{i, \mathtxt{OBS}}^2.
  \label{eq:app:b:sigma_sum}
\end{equation}
The specified error ranges for the \ac{MMM} approaches are calculated
similarly: The constant \ac{SPE} per grid cell is estimated by the mean
\ac{RMSEP} given in the leave-one-model-out \ac{CV} experiment (see
\cref{subsec:02:model_weighting}). For the plain \ac{MMM}, this is the only
source of uncertainty. For the rescaled \ac{MMM}, the total error can be
calculated similarly to \cref{eq:app:b:sigma_sum} without the last term
(observational uncertainty).

@@

A convenient way to gain information about statistical models is to analyze the
residuals $\varepsilon_i$, which are defined as the difference between the true
value of the target variable $y_i$ and the predicted value $\hat{y}_i$ at a
sample point $i$ with known ground truth:
\begin{equation}
  \varepsilon_i = y_i - \hat{y}_i.
  \label{eq:app:b:residuals}
\end{equation}
A common way to visualize the residuals is to plot their \acfp{PDF} (see
\cref{fig:app:b:residuals}). The two panels (for steps 2a and 2b) show that the
\ac{ML} model is not overfitting in both cases: The distributions of the
training data and the independent hold-out data are very similar. Moreover, the
\acp{PDF} do not show significant biases, as the residuals are approximately
unbiased (zero mean) for the training and the test dataset. This justifies the
use of the \ac{RMSEP} to estimate the \ac{SPE} since for unbiased residuals
the \ac{RMSEP} is equal to the standard deviation of the residuals.

@@

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4a.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4b.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:b}
  \end{subfigure}
  \\
    \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4c.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4d.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:d}
  \end{subfigure}
  \\
    \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4e.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4f.pdf}
    \caption{}
    \label{fig:app:b:step2a_results:f}
  \end{subfigure}
  \caption[
    Geographical distributions of the absolute \acf{GPP} at the end of the
    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
    statistical models.
  ]{
    Geographical distributions of the absolute \acf{GPP} at the end of the
    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
    statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
    \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO} model using only the
    historical \acs{GPP} as a single predictor. (d) \Acf{GBRT} model using only
    the historical \acs{GPP} as a single predictor. (e) \acs{LASSO} model using
    all predictors. (f) \acs{GBRT} model using all predictors.
    \AdaptedFrom{Schlund2020}.
  }
  \label{fig:app:b:step2a_results}
\end{figure}

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5a.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5b.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5c.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5d.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:d}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5e.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5f.pdf}
    \caption{}
    \label{fig:app:b:step2a_results_errors:f}
  \end{subfigure}
  \caption[
    Geographical distributions of the \aclp{SPE} of the absolute \acf{GPP} at
    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
    different statistical models.
  ]{
    Geographical distributions of the \acfp{SPE} of the absolute \acf{GPP} at
    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
    different statistical models. Details on the calculation of the \acs{SPE}
    are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5 \acf{MMM}.
    (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO}
    model using only the historical \acs{GPP} as a single predictor. (d)
    \Acf{GBRT} model using only the historical \acs{GPP} as a single predictor.
    (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT} model using all
    predictors. The \acs{SPE} is minimal for the \acs{GBRT} model using all
    predictors. \AdaptedFrom{Schlund2020}.
  }
  \label{fig:app:b:step2a_results_errors}
\end{figure}
